Cluster Difference Imaging Photometric Survey. III. An Open Cluster With a 300\,pc Halo

### 2021-03-02 cluster_paper/ms.tex
update manuscript

%TODO:
%
% n rotation period logs: 2238 % ~/proj/cdips/results/allvariability_reports/NGC_2516/logs
% ... which came from ~/Dropbox/proj/cdips/data/cluster_data/cdips_catalog_split/OC_MG_FINAL_v0.4_publishable_CUT_NGC_2516.csv
% n all variability reports: 2201 % ~/proj/cdips/results/allvariability_reports/NGC_2516/reports

@@

%\newcommand{\nautorot}{1{,}144\ } % % For nequal<=0, nclose<=100, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt

\newcommand{\nkinematic}{3{,}018\ } % "fullfaint" kinematic sample, core+halo. from any call of earhat.helpers._get_fullfaint_dataframes
\newcommand{\ncdips}{2,238\ } % e.g., from OC_MG_FINAL_v0.4_publishable_CUT_NGC_2516.csv. checked with plot_target_vs_actual_list_sanity_checks.py

\newcommand{\ncalibration}{9{,}619{,}784\ } % /nfs/phtess2/ar0/TESS/PROJ/lbouma/CDIPS_LCS/counts/check_sum, 2021.02.02
\newcommand{\nnbhd}{13{,}843\ } % make_compstar_NGC_2516_sourcelist.py
\newcommand{\nnbhdcaliblc}{9{,}790\ } % get_auto_rotation_periods.py
\newcommand{\nnbhdcalibstar}{1{,}987\ } % get_auto_rotation_periods.py

\newcommand{\nautorotdenominator}{1{,}641\ } % plot_autorotation.py, w/ runid=NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
\newcommand{\nautorotnumerator}{987\ } % plot_autorotation.py, For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
\newcommand{\ncompstardenominator}{820\ } % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
\newcommand{\ncompstarnumerator}{365\ } % plot_autorotation.py
\newcommand{\ncompfrac}{44.5\%} % plot_autorotation.py
\newcommand{\nautovscompstardenominator}{705\ } % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
\newcommand{\nautovscompstarnumerator}{374\ } % plot_autorotation.py
\newcommand{\nautofrac}{53.1\%} % plot_autorotation.py

@@

% Context, aims, methods, results, conclusions.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters. 
  The stars in these diffuse halos could be the tidal escapees of the
  cluster cores, or they could be remnants of an initially dispersed
  star formation complex.
  They could also be field stars---the false positives of overly
  aggressive clustering algorithms.
  In this study, we focus on a halo with a length of 300\,pc reported
  around the $\approx$175\,Myr open cluster NGC\,2516.
  %TODO: says who?
  The classical tidal radius of the cluster is $\approx$10\,pc.
  Combining photometry from Gaia, rotation periods from TESS, and
  lithium measurements from Gaia-ESO and GALAH, we find that the halo
  of NGC\,2516 is real and coeval with the core.
  Two in three kinematically selected halo members 25-250\,pc from the
  cluster center show rotation periods consistent with a
  gyrochronological age of 120\,Myr.
  A comparison sample of field stars shows no such trend.
  The rotation period and lithium abundances of the cluster stars are
  also anti-correlated at fixed stellar mass, as has been observed in
  similar stellar populations.
  This work expands the set of confirmed NGC\,2516 members by a factor
  of $\approx$2, and quantifies the degree to which Gaia-based
  analyses can identify dispersed stellar populations against a
  background of field star contaminants.
  We highlight the implications for spectroscopic survey targeting,
  open cluster dispersal, and planet searches around young stars.
\end{abstract}

@@

Star clusters form in hierarchically structured molecular clouds
\citep{shu_star_1987}.  Over the first 10\,Myr, ``feedback'' effects
including protostellar outflows, photoionization, radiation pressure,
and supernova shocks disperse the gas out of the cloud
\citep{krumholz_star_2019}.  Since only a small fraction of the cloud
mass is converted into stars ($\sim 1\%$), gas dispersal enables the
majority ($\sim$90\%) of stars in the cluster to escape the cluster's
gravitational well \citep{lada_embedded_2003}.

@@

Finding the stars that have dispersed from their natal clusters into
the galactic field is of interest for a few reasons.  One is to
understand the predominant conditions under which stars and star
clusters form \citep[{\it e.g.},][]{wright_kinematics_2018}.  Another
is to understand the Sun's birth environment \citep{adams_birth_2010}.
Did the Sun form in an open cluster, and if so, what type? Is there
any hope at identifying the stars that formed near the Sun?  Open
clusters and their dispersed remnants also provide a test case for
``chemical tagging'' \citep[{\it
e.g.},]{freeman_new_2002,hogg_chemical_2016,ness_dopplergangers_2018}.
The converse holds too: one could use chemical tagging to test the
viability of kinematic clustering methods.

@@

The detection of dispersed stellar associations, although possible
before Gaia, has now reached a breakthrough pace \citep[{\it
e.g.},][]{bergond_gravitational_2001,zuckerman_young_2004,oh_comoving_2017,cantatgaudin_gaia_2018,gagne_banyan_XII_2018,gagne_banyan_XIII_2018,kounkel_apogee2_2018,zari_3d_2018,kounkel_untangling_2019,furnkranz_comabernbhr_2019}.
The resuls can be summarized as follows. On one end are coeval
populations with no discernable cores ({\it e.g.}, Psc Eri and $\mu$
Tau,
\citealt{meingast_psceri_2019,curtis_tess_2019,gagne_mutau_2020}).  On
the other are hierarchically structured populations with many over and
under-densities  \citep[{\it e.g.} the Sco-Cen and Vela OB2
associations][]{pecaut_star_2016,cantatgaudin_velaOB2_2019}.  Here, we
focus on the middle: low-density halos associated with a single
overdensity, typically an open cluster that was known before Gaia
\citep[see][]{kounkel_untangling_2019,kounkel_untanglingII_2020,meingast_2021}.
In some cases, these halos could correspond to tidal tails, as have
been reported in the Hyades
\citep{meingast_hyades_2019,roser_hyades_2019}, the Ursa Major moving
group \citep{gagne_lowmassUMA_2020}, and in Coma Berenices
\citep{tang_comaber_2019}.

One point of difficulty however is that different clustering methods
yield differing levels of both sensitivity and precision
\citep{hunt_clustering_2020}.  Using say a Gaussian Mixture Model
tautologically yields open clusters that are elliptical \citep[{\it
e.g.},][]{wallace_m4_2018}.  Some unsupervised methods yield dispersed
and asymmetric structures with number densities down to 100 times less
than the field around the same regions (see {\it e.g.}
\citealt{kounkel_untangling_2019}, hereafter
\citetalias{kounkel_untangling_2019}], and \citealt{meingast_2021},
hereafter \citetalias{meingast_2021}).  Discrepancies between
different clustering methods leads to the question of how contaminated
the resulting samples are, and in some cases whether the reported
structures are truly coeval stellar populations.

We have recently been making TESS light curves of stars reported to be
members of such populations (CDIPS, \citealt{bouma_cdipsI_2019}).  Our
analysis of TESS Cycle 1 (Sectors 1-13) yielded light curves of
\numsouthernuniqlcs candidate cluster members in the Southern Ecliptic
hemisphere, available on
MAST\footnote{\url{https://archive.stsci.edu/hlsp/cdips}}.

As part of this broader project, we focus here on a modest question:
in just a single rich southern open cluster, is the cluster halo truly
coeval with the core?  We chose NGC\,2516 for this analysis since it
is young ($\approx$120\,Myr) and sufficiently nearby
($d\approx400$\,pc) to facilitate photometric rotation measurements
using TESS, and some degree of spectroscopic analysis.  Its halo is
decidedly non-spherical: \citetalias{kounkel_untangling_2019} reported
its dimensions as $\approx20\times10\times350$\,pc.  We want to know:
is the halo real? To what extent can we use Gaia alone to reliably
identify age-dated needles in the haystack of field stars?  What are
the implications, observationally and theoretically, if we can
identify the dispersed halos of open clusters?

A brief note on terminology.  Low-density stellar associations
connected to a dense population (the ``core'') have been described as
being in ``halos'', ``strings'', ``coronae'', ``snakes'', ``the
outskirts'', and ``tidal tails'' \citep[{\it
e.g.},][]{davenport_death_2010,kounkel_untangling_2019,roser_hyades_2019,tian_discovery_2020,meingast_2021}.
The latter term implies a particular model for the formation of the
association.  We opt to use ``halo'' because it is concise and model
agnostic.  It suffers however from the suggestion of spherical
symmetry.  The halo of NGC\,2516 is anything but.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf The dense and diffuse components of NGC\,2516 across position and
  velocity space.} The core (black) was identified by
  \citet{cantatgaudin_gaia_2018} using Gaia DR2, and is coincident with
  the visual overdensity of stars canonically accepted as ``the
  cluster''.  The halo (blue) was identified by
  \citet{kounkel_untangling_2019} and \citet{meingast_2021} using less restrictive membership
  assignment algorithms described in Appendix~\ref{app:clustering}.  The field
  sample is randomly drawn from a $(\alpha, \delta, \pi)$ cube
  centered on the cluster.  The low volume density of the
  halo stars makes it difficult to visually distinguish
  the field and the halo samples.
  \label{fig:gaia6d}
	}
\end{figure*}

We selected candidate \cn\ members based on those available in the
literature.  While a number of pre-Gaia catalogs were available, the
purity and accuracy of the Gaia-derived results are the current state
of the art.
We therefore adopted what we viewed as the most interesting clustering
samples to compare: those of \citet{cantatgaudin_gaia_2018}
(\citetalias{cantatgaudin_gaia_2018}),
\citet{kounkel_untangling_2019}
(\citetalias{kounkel_untangling_2019}),
and \citet{meingast_2021} (\citetalias{meingast_2021}).
A useful visualization of these samples is available online\footnote{
  \url{https://homepage.univie.ac.at/stefan.meingast/coronae.html},
  made by \citet{meingast_2021}.  }.
While we could have performed our own clustering analysis based on the
Gaia data, such an effort would be only replicating the work of these
investigators.  We opt to instead use their studies as starting
points.

%FIXME FIXME FIXME FROM HERE

% Theia 613
\citetalias{cantatgaudin_gaia_2018} reported 1106 candidate members of
\cn, brighter than $G=18$\,mag.  \citet{kounkel_untangling_2019}
reported 3003 candidate members, and included stars up to
$\approx$1\,mag fainter.  
Finally, 1577 of the 1860 \citet{meingast_2021} sources (85\%)
were included in the \citet{kounkel_untangling_2019} sample.  
Each study applied unsupervised clustering techniques
to the second Gaia Data Release, as detailed in
Appendix~\ref{app:clustering}.

%FIXME 
\citetalias{cantatgaudin_gaia_2018} considered a cone of radius {\bf
XX DEGREES} centered on the cluster's position.
Given the classically reported tidal radius of {\it e.g.,} 9\,pc, CITE
Piskunov+2008, which is 8$^\circ$ at the cluster distance of
$\approx$400\,pc, perhaps they should have considered stars further
out.

%TODO calculate these physical sizes carefully
Figure~\ref{fig:gaia6d} shows the cluster members reported by each
study.  The \citetalias{cantatgaudin_gaia_2018} members are all within
a few degrees of the cluster center, while the
\citetalias{kounkel_untangling_2019} members span tens of degrees.  In
physical units, this corresponds to a size difference of XX\,pc versus
YY\,pc.  This difference is somewhat tautological, because
\citetalias{cantatgaudin_gaia_2018} did not extend their search for
members out to tens of degrees from the cluster center, since their
clustering algorithm relied on the contrast between cluster and field
stars in the astrometric space of ($\mu_{\alpha'}, \mu_\delta, \pi$).
Nonetheless, the membership catalog of \citetalias{cantatgaudin_gaia_2018}
echoes that of many previous investigators (CITE, CITE, CITE), and is
consistent with the general {\it visual} impression that one has when
looking at \cn\ visually: it seems to be at most {\bf X$^\circ$}, and
not a big stream.
This discrepancy begs the question.  What is the true structure of
\cn?  Are the core and halo truly coeval?

To answer these questions, we adopt the
\citetalias{cantatgaudin_gaia_2018} stars as candidate ``core''
members and the \citetalias{kounkel_untangling_2019} stars as
candidate ``halo'' members.  If a star is in both lists, it is a
``core'' member.  1091 of the 1106 core members are in both lists.
This leaves 1912 candidate members in the halo.
Concatenating the two lists yields \nkinematic candidate members.

In this work, we also define a set of nearby field stars in the
``neighborhood'' of \cn.  Based on the observed distribution of halo
members, we draw these stars randomly from the following ``cube'' in
right ascension, declination, and parallax:
\begin{align}
  \alpha\,[\mathrm{deg}] &\in [108, 132], \\
  \delta\,[\mathrm{deg}] &\in [-76, -45], \\
  \pi\,[\mathrm{mas}] &\in [1.5, 4.0].
\end{align}
We imposed a magnitude limit of $G=19$\,mag, and ran the queries
using the \texttt{astroquery.gaia} module (CITE).
% TODO: why?
We allowed the number of stars in the comparison sample to exceed that
in the cluster sample by a factor of $\approx$5, to ensure broad sampling
of stellar masses and evolutionary states.
We also required the comparison sample to not overlap with the cluster
sample, which led to the omission of 1.1\% of the stars drawn over the
volume noted above.

@@

The first check on the plausibility of the candidate cluster members
being coeval was already performed by
\citetalias{cantatgaudin_gaia_2018},
\citetalias{kounkel_untangling_2019}, and \citetalias{meingast_2021}.
That check is to see whether the HR diagrams of the cluster members
support the claim that they are coeval.

Figure~\ref{fig:hr} presents similar results to what these
investigators have already found.  The core members of the cluster
show a clean sequence consistent with stars with a fixed age and
metallicity, and varying mass.  The halo members show a roughly
consistent main-sequence, with somewhat greater scatter.  One
possible explanation for the additional scatter could be that the
halo is more contaminated by field stars.  For instance, $\approx5$
evolved halo stars are visible in the field-star portion of the red
giant branch (RGB).  Given the age of the system, these stars must be
field interlopers.

A separate possible explanation for scatter in the halo's HR diagram
could be differential reddening across different sightlines.  The halo
is reported to span 20$^\circ$ on-sky, and varies in position from
about $b=-12^\circ$ to $b=-20^\circ$, with the stars closest to the
galactic plane also being further from the Sun by up to 200\,pc
(Figure~\ref{fig:gaia6d}).  Some empirical evidence for differential
reddening is discussed further in Appendix~\ref{app:diffred}. Based on
the available evidence, we expect that both field star contamination
and differential reddening could play a role in the larger scatter of
the halo relative to the core.  A third possibility that we have not
explored is whether the binary fraction could also differ between the
different regions of the cluster.

In the right panel of Figure~\ref{fig:hr}, we compare the observed
photometry with PARSEC isochrones
\citep{bressan_parsec_2012,chen_improving_2014,chen_parsec_2015,marigo_new_2017}.
We used the web
interface\footnote{\url{http://stev.oapd.inaf.it/cgi-bin/cmd},
\texttt{2021-02-26}, \texttt{CMD3.4}, \texttt{YBC} bolometric
corrections as in \citet{chen_2019_YBC}.} to interpolate these
isochrones at $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$, and applied
the reddening correction described in Section~\ref{subsec:tess}.  We
assumed a reddening of $E(B-V)=0.102\pm0.019$, $R_V=3.1$, and applyied
the \citet{odonnell_1994} SED-dependent extinction law star-by-star
through the PARSEC interface.  For the metallicity, we considered a
range of super and sub-solar metallicities to fit as much of the locus
from $0.5<Bp-Rp<1.5$ as possible, and settled on the slightly
super-solar $[M/H]=0.06$.  Sub-solar metallicities led to model
predictions too blue along the main sequence by $\approx$0.1\,mag.
Our adopted metallicity agrees with that found by
\citet[][Sec~4.4.4]{cummings_2011_li_iron}, and with the iron
abundance recently determined by the Gaia-ESO team
\citep{baratella_gaiaeso_2020}, which represented an up-revision from
an earlier sub-solar metallicity \citep{randich_gaiaeso_2018}.  While
\citet{bailey_rv_2018} report a slightly sub-solar metallicity for
the cluster, we prefer super-solar based on the goodness of fit from
the photometry.

@@

Our primary goal is to determine whether the ages of the core and halo
are consistent.  The left panel of Figure~\ref{fig:hr} shows
empirically that they are: stars past the turnoff in both the halo and
core are on a consistent locus.  The right panel of
Figure~\ref{fig:hr} demonstrates the precision with which the claim can be
made.  The MSTO stars are consistent with the 178 and 316 Myr models,
but are bluer than the 100 Myr model.  The RGB stars are consistent
only with the 178 Myr ($\log_{10} t/{\rm yr} = 8.25$) model.  Based on
the assumption that the width of the MSTO can be attributed to binary
stars, we are most interested in the blue edge. This appears most
compatible with the 178 Myr model.  These results are consistent with
the MSTO age range of 165 to 195\,Myr found by \citet{cummings_2018},
and we refer to their work for a more precise model-dependent
comparison.

>>>>

@@

\newcommand{\nkinematic}{3{,}298\ } % "fullfaint" kinematic sample, core+halo (CG18+KC19+M21). from any call of earhat.helpers._get_fullfaint_dataframes
\newcommand{\nnbhd}{13{,}843\ } % make_compstar_NGC_2516_sourcelist.py
\newcommand{\ncore}{1{,}106}  % "fullfaint" kinematic sample, CG18. from any call of earhat.helpers._get_fullfaint_dataframes
\newcommand{\nhalo}{2{,}2192} % "fullfaint" kinematic sample, KC19+M21. from any call of earhat.helpers._get_fullfaint_dataframes

% n rotation period logs: 2238 % ~/proj/cdips/results/allvariability_reports/NGC_2516/logs
% ... which came from ~/Dropbox/proj/cdips/data/cluster_data/cdips_catalog_split/OC_MG_FINAL_v0.4_publishable_CUT_NGC_2516.csv
% n all variability reports: 2201 % ~/proj/cdips/results/allvariability_reports/NGC_2516/reports
\newcommand{\ncdips}{2,238\ } % e.g., from OC_MG_FINAL_v0.4_publishable_CUT_NGC_2516.csv. checked with plot_target_vs_actual_list_sanity_checks.py

\newcommand{\ncalibration}{9{,}619{,}784\ } % /nfs/phtess2/ar0/TESS/PROJ/lbouma/CDIPS_LCS/counts/check_sum, 2021.02.02
\newcommand{\nnbhdcaliblc}{9{,}790\ } % get_auto_rotation_periods.py
\newcommand{\nnbhdcalibstar}{1{,}987\ } % get_auto_rotation_periods.py

%FIXME these change depending on your cutoff. ensure the final numbers agree!
\newcommand{\nautorotdenominator}{1{,}641\ } % plot_autorotation.py, w/ runid=NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
\newcommand{\nautorotnumerator}{987\ } % plot_autorotation.py, For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
\newcommand{\ncompstardenominator}{820\ } % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
\newcommand{\ncompstarnumerator}{365\ } % plot_autorotation.py
\newcommand{\ncompfrac}{44.5\%} % plot_autorotation.py
\newcommand{\nautovscompstardenominator}{705\ } % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
\newcommand{\nautovscompstarnumerator}{374\ } % plot_autorotation.py
\newcommand{\nautofrac}{53.1\%} % plot_autorotation.py

@@

% Context, aims, methods, results, conclusions.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters. 
  The stars in these diffuse halos could be the tidal escapees of the
  cluster cores, or they could be remnants of an initially dispersed
  star formation complex.
  They could also be field stars---the false positives of overly
  aggressive clustering algorithms.
  In this study, we focus on a halo with a length of 300\,pc reported
  around the $\approx$175\,Myr open cluster NGC\,2516.
  %TODO: Says who? Says Piskunov+08
  The classical tidal radius of the cluster is $\approx$10\,pc.
  Combining photometry from Gaia, rotation periods from TESS, and
  lithium measurements from Gaia-ESO and GALAH, we find that the halo
  of NGC\,2516 is real and coeval with the core.
  Two in three kinematically selected halo members 25-250\,pc from the
  cluster center show rotation periods consistent with a
  gyrochronological age of 120\,Myr.
  A comparison sample of field stars shows no such trend.
  The rotation period and lithium abundances of the cluster stars are
  also anti-correlated at fixed stellar mass, as has been observed in
  similar stellar populations.
  This work expands the set of confirmed NGC\,2516 members by a factor
  of $\approx$2, and quantifies the degree to which Gaia-based
  analyses can identify dispersed stellar populations against a
  background of field star contaminants.
  We highlight the implications for spectroscopic survey targeting,
  open cluster dispersal, and planet searches around young stars.
\end{abstract}

@@

Star clusters form in hierarchically structured molecular clouds
\citep{shu_star_1987}.  Over the first 10\,Myr, feedback effects
including protostellar outflows, photoionization, radiation pressure,
and supernova shocks disperse the gas out of the cloud
\citep{krumholz_star_2019}.  Since only a small fraction of the cloud
mass is converted into stars ($\sim 1\%$), gas dispersal enables the
majority ($\sim$90\%) of stars in the cluster to escape the cluster's
gravitational well \citep{lada_embedded_2003}.

@@

% NOTE: words in this paragraph need work
Finding the stars that have dispersed from their natal clusters into
the galactic field is of interest for a few reasons.  One is to
understand the predominant conditions under which stars and star
clusters form \citep[{\it e.g.},][]{wright_kinematics_2018}.  Another
is to understand the Sun's birth environment \citep{adams_birth_2010}.
Did the Sun form in an open cluster, and if so, what type? Is there
any hope at identifying the stars that formed near the Sun?  Open
clusters and their dispersed remnants also provide a test case for
``chemical tagging'' \citep[{\it
e.g.},]{freeman_new_2002,hogg_chemical_2016,ness_dopplergangers_2018}.
The converse holds too: one can use chemical tagging to test the
viability of kinematic clustering methods.

@@

The detection of dispersed stellar associations, although possible
before Gaia, has now reached a breakthrough pace \citep[{\it
e.g.},][]{bergond_gravitational_2001,zuckerman_young_2004,oh_comoving_2017,cantatgaudin_gaia_2018,gagne_banyan_XII_2018,gagne_banyan_XIII_2018,kounkel_apogee2_2018,zari_3d_2018,kounkel_untangling_2019,furnkranz_comabernbhr_2019}.
The populations found by these studies can be summarized as follows.
On one end are coeval groups with no discernable cores ({\it e.g.},
Psc Eri and $\mu$ Tau,
\citealt{meingast_psceri_2019,curtis_tess_2019,gagne_mutau_2020}).  On
the other are hierarchically structured associations with many over
and under-densities  \citep[{\it e.g.} Sco-Cen and Vela
OB2][]{pecaut_star_2016,cantatgaudin_velaOB2_2019}.  Here, we focus on
an intermediate regime: low-density halos associated with a single
overdensity, typically an open cluster that was known before Gaia
\citep[see][]{kounkel_untangling_2019,kounkel_untanglingII_2020,meingast_2021}.
In some cases, these halos could correspond to tidal tails, as have
been reported in the Hyades
\citep{meingast_hyades_2019,roser_hyades_2019}, the Ursa Major moving
group \citep{gagne_lowmassUMA_2020}, and in Coma Berenices
\citep{tang_comaber_2019}.

One point of difficulty however is that different clustering methods
yield differing levels of both sensitivity and precision
\citep{hunt_clustering_2020}.  Using say a Gaussian Mixture Model
tautologically yields open clusters that are elliptical \citep[{\it
e.g.},][]{wallace_m4_2018}.  Some unsupervised methods yield dispersed
and asymmetric structures with number densities down to 100 times less
than the field around the same regions (see {\it e.g.}
\citealt{kounkel_untangling_2019} and \citealt{meingast_2021}).
Difference between different clustering methods raises the issue of
how contaminated the resulting samples are, and in some cases whether
the reported structures are truly coeval stellar populations at all.

We have recently been making TESS light curves of stars reported to be
members of coeval populations (CDIPS, \citealt{bouma_cdipsI_2019}).
Our analysis of TESS Cycle 1 (Sectors 1-13) yielded light curves of
\numsouthernuniqlcs candidate cluster members in the Southern Ecliptic
hemisphere, which are available on
MAST\footnote{\url{https://archive.stsci.edu/hlsp/cdips}}.

As part of this broader project, we focus here on a modest question:
in just a single rich southern open cluster, is the cluster halo truly
coeval with the core?  We chose NGC\,2516 for this analysis since it
is young (100--200\,Myr) and sufficiently nearby ($d\approx400$\,pc)
to facilitate photometric rotation measurements using TESS, and some
degree of spectroscopic analysis.  Its halo is decidedly
non-spherical: \citet{kounkel_untangling_2019} reported its dimensions
as $\approx20\times10\times350$\,pc.  We want to know: is the halo
real? To what extent can we use Gaia alone to reliably identify
age-dated needles in the haystack of field stars?  What are the
implications, observationally and theoretically, if we can identify
the dispersed halos of open clusters?

%TODO: double check this owrding.
A brief note on terminology.  Low-density stellar associations
connected to a dense population (the ``core'') have been described as
being in ``halos'', ``strings'', ``coronae'', ``snakes'',
``outskirts'', and ``tidal tails'' \citep[{\it
e.g.},][]{davenport_death_2010,kounkel_untangling_2019,roser_hyades_2019,tian_discovery_2020,meingast_2021}.
The latter term implies a particular model for the formation of the
association.  We opt to use ``halo'' because it is concise and model
agnostic.  It suffers however from the suggestion of spherical
symmetry.  The halo of NGC\,2516 might be better-described as ``tidal
tails'', but the primary goal of this study however is
not to verify this claim.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf The dense and diffuse components of NGC\,2516 across
  position and velocity space.} The core (black) was analyzed by
  \citet{cantatgaudin_gaia_2018} using Gaia DR2, and is coincident
  with the visual overdensity of stars canonically accepted as ``the
  cluster''.  The halo (blue) is a concatenation of the studies by
  \citet{kounkel_untangling_2019} and \citet{meingast_2021}, which
  used less restrictive membership assignment algorithms described in
  Appendix~\ref{app:clustering}.  The field sample is randomly drawn
  from a $(\alpha, \delta, \pi)$ cube centered on the cluster.  The
  low volume density of the halo stars makes it difficult to visually
  distinguish the field and the halo samples.
  \label{fig:gaia6d}
	}
\end{figure*}

We selected candidate \cn\ members based on those available in the
literature.  While a number of pre-Gaia catalogs were available, the
purity and accuracy of the Gaia-derived results are the current state
of the art.  We therefore adopted what we viewed as the most
interesting clustering samples to compare: those of
\citet{cantatgaudin_gaia_2018}, \citet{kounkel_untangling_2019} and
\citet{meingast_2021}.  We refer to these studies in the following as
\citetalias{cantatgaudin_gaia_2018},
\citetalias{kounkel_untangling_2019}, and \citetalias{meingast_2021}
respectively.  A useful visualization of the different samples is
available online\footnote{
  \url{https://homepage.univie.ac.at/stefan.meingast/coronae.html},
  made by \citet{meingast_2021}.}.  While we considering performing
our own clustering analysis using the Gaia data, such an effort would
in effect only replicate the work of these investigators.  We opt to
instead use their studies as starting points.

The Gaia clustering studies each used slightly different selection
functions, with different results.
\citetalias{cantatgaudin_gaia_2018} considered stars brighter than
$G=18$\,mag within a few degrees of the cluster center, and reported
1106 candidate members of \cn.  \citetalias{kounkel_untangling_2019}
and \citetalias{meingast_2021} considered stars up to $\approx$1\,mag
fainter, and reported 3003 and 1860 members respectively.  The
unsupervised clustering techniques that each of these studies applied
to the second Gaia Data Release are discussed
Appendix~\ref{app:clustering}, as is the overlap between their
resulting membership samples.

% %FIXME 
% \citetalias{cantatgaudin_gaia_2018} considered a cone of radius {\bf
% XX DEGREES} centered on the cluster's position.
% Given the classically reported tidal radius of {\it e.g.,} 9\,pc, CITE
% Piskunov+2008, which is 8$^\circ$ at the cluster distance of
% $\approx$400\,pc, perhaps they should have considered stars further
% out.

Figure~\ref{fig:gaia6d} shows the cluster members reported by each
study.  The \citetalias{cantatgaudin_gaia_2018} members are all within
a few degrees of the cluster center, while the
\citetalias{kounkel_untangling_2019} and \citetalias{meingast_2021}
members span tens of degrees.  In physical units, the physical size of
the entire structure is $\approx20\times10\times350$\,pc.
As a side note, stars in the upper right quadrant of $\mu_{\alpha'}$
{\it vs.} $\mu_\delta$ were reported almost entirely by
\citet{meingast_2021}.

In Figure~\ref{fig:gaia6d}, and for the remainder of the study, we
describe the \citetalias{cantatgaudin_gaia_2018} members as the
``core'' of the cluster, and any non-overlapping
\citetalias{kounkel_untangling_2019} and \citetalias{meingast_2021}
members as the ``halo''.  These definitions imply that there are
\ncore core members, and \nhalo halo members.  This distinction is
perhaps tautological, because \citetalias{cantatgaudin_gaia_2018} did
not extend their search for members out to tens of degrees from the
cluster center.  Nonetheless, the membership catalog of
\citetalias{cantatgaudin_gaia_2018} echoes that of many previous
investigators (CITE, CITE, CITE), and is consistent with the general
{\it visual} impression that one has when looking at \cn\ in the night
sky: it appears to span $\approx$2$^\circ$, and not $\approx20^\circ$.

The discrepancy between our canonical understanding of open clusters
and the possibility of a diffuse population existing around \cn\
prompted our initial interest in this analysis.  What is the true
structure of \cn?  Are the core and halo truly coeval?

Outside of the core and halo, we also define a set of nearby field
stars in the ``neighborhood'' of \cn.  Based on the observed
distribution of halo members, we drew these stars randomly from the
following ``cube'' in right ascension, declination, and parallax:
\begin{align}
  \alpha\,[\mathrm{deg}] &\in [108, 132], \\
  \delta\,[\mathrm{deg}] &\in [-76, -45], \\
  \pi\,[\mathrm{mas}] &\in [1.5, 4.0].
\end{align}
We imposed a magnitude limit of $G=19$\,mag, and ran the queries using
the \texttt{astroquery.gaia} module (CITE).  We allowed the number of
stars in the comparison sample to exceed that in the cluster sample by
a factor of $\approx$5, to ensure broad sampling of stellar masses and
evolutionary states.  We also required the comparison sample to not
overlap with the cluster sample, which led to the omission of 1.1\% of
the stars drawn from the volume noted above.

@@

The first check on the plausibility of the candidate cluster members
being coeval is to analyze the Gaia Hertzsprung-Russell (HR) diagrams.  Comparable analyses
have already been performed by \citetalias{cantatgaudin_gaia_2018},
\citetalias{kounkel_untangling_2019}, and \citetalias{meingast_2021}.

Figure~\ref{fig:hr} shows the HR diagram in the space of absolute Gaia
$G$ magnitude as a function of observed $Bp-Rp$ color.  The core
members of the cluster show a clean sequence consistent with stars
with a fixed age and metallicity, and varying mass.  The halo members
mostly show a consistent main-sequence that has somewhat greater
scatter.  One possible explanation for the additional scatter could be
that the halo is more contaminated by field stars.  For instance,
$\approx5$ evolved halo stars are visible in the field-star portion of
the red giant branch (RGB).  Given the age of the system, these stars
must be field interlopers.

A separate possible explanation for scatter in the halo's HR diagram
could be differential reddening across different sightlines.  The
reported halo spans 20$^\circ$ on-sky, and varies in position from
about $b=-12^\circ$ to $b=-20^\circ$, with the stars closest to the
galactic plane also being further from the Sun by up to 200\,pc
(Figure~\ref{fig:gaia6d}).  Some empirical evidence for differential
reddening is discussed in Appendix~\ref{app:diffred}. Based on the
available evidence, we expect that both field star contamination and
differential reddening could play a role in the larger scatter of the
halo relative to the core.  A third possibility that we have not
explored is whether the binary fraction could also differ between the
different regions of the cluster.

In the right panel of Figure~\ref{fig:hr}, we compare the observed
Gaia EDR3 photometry with PARSEC isochrones
\citep{bressan_parsec_2012,chen_improving_2014,chen_parsec_2015,marigo_new_2017}.
We used the web
interface\footnote{\url{http://stev.oapd.inaf.it/cgi-bin/cmd},
\texttt{2021-02-26}, \texttt{CMD3.4}, \texttt{YBC} bolometric
corrections as in \citet{chen_2019_YBC}.} to interpolate these
isochrones at $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$, and applied
the reddening correction described in Section~\ref{subsec:tess}.  We
assumed a reddening of $E(B-V)=0.102\pm0.019$, $R_V=3.1$, and applied
the \citet{odonnell_1994} SED-dependent extinction law star-by-star
through the PARSEC interface.  For the metallicity, we considered a
range of super and sub-solar metallicities to fit as much of the locus
from $0.5<Bp-Rp<1.5$ as possible, and settled on the slightly
super-solar $[M/H]=0.06$.  Sub-solar metallicities led to model
predictions too blue along the main sequence by $\approx$0.1\,mag.
Our adopted metallicity agrees with that found by
\citet[][Sec~4.4.4]{cummings_2011_li_iron}, and with the iron
abundance recently determined by the Gaia-ESO team
\citep{baratella_gaiaeso_2020}, which represented an up-revision from
an earlier sub-solar metallicity \citep{randich_gaiaeso_2018}.  While
\citet{bailey_rv_2018} report a slightly sub-solar metallicity for the
cluster, we prefer super-solar based on the goodness of fit from the
photometry.

@@

Our primary goal is to determine whether the ages of the core and halo
are consistent.  The left panel of Figure~\ref{fig:hr} suggests that
they are: stars past the turnoff in both the halo and core are on a
consistent locus.  The right panel of Figure~\ref{fig:hr} demonstrates
the precision with which the claim can be made.  The MSTO stars are
consistent with the 178 and 316 Myr models, but are bluer than the 100
Myr model.  The RGB stars are consistent only with the 178 Myr
($\log_{10} t/{\rm yr} = 8.25$) model.  Based on the assumption that
the width of the MSTO can be attributed to binary stars, we are most
interested in the blue edge. This appears most compatible with the 178
Myr model.  These results are consistent with the MSTO age range of
165 to 195\,Myr found by \citet{cummings_2018}, and we refer to their
work for a more precise model-dependent comparison.

% FIXME: do you want to do the procedure below? Seems like overkill
% given the Cummings study.
